An Expressive and Efficient Solution to the Service Selection Problem

### 2010-06-29 publicaciones/iswc10/paperISWC10-v3.tex
typos corregidos en paperISWC10-v3

Existing Web infrastructures support the publication and access to a
tremendous amount of Web data sources, some of which that can be labeled
and converted into Semantic Web Services by using existing annotation
tools like the one proposed by Ambite et al.\ \cite{AmbiteISWC09}.
Once a large dataset of Semantic Web Services become available, users
will require techniques to effectively select the services that meet
their requirements.
In order to achieve this goal, the services in the dataset must be
tagged with their functional and non-functional properties, and
the users' preferences and requirements must be formally described
as well.
In this paper, we extend an approach traditionally used in the area of
data integration to solve the problem of selecting the best services
that meet a user request, a problem that we call in this paper the
Service Selection Problem (SSP).

@@

As it is shown below, LAV descriptions are achieved with mappings
that define the services as conjunctive queries involving the generic
concepts in the ontology. Thus, every time that a service changes or
a new one becomes available, only a tiny fraction of the mappings
must be updated, usually just one mapping.
Likewise, user requests can be modeled as conjunctive queries over the
generic concepts in a way that the SSP can be cast as the problem
of rewriting a query in terms of a set of views, the so-called Query
Rewriting Problem (QRP) that is well-known in the area of data
integration~\cite{Chen05,JaudoinPRST05}, query optimization and data
maintenance~\cite{AfratiLU07,levy:bucket}, and for which several
scalable approaches had been
proposed~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
Furthermore, users' preferences and constraints on the set of the
selected services to serve a given request are supported by the
adoption of a simple yet expressive language for preferences. 
These preferences and constraints further refine and rank the
the set of valid rewritings of the posed query, in a way that
the best solution to the SSP correspond to the best-ranked 
valid rewritings of the corresponding QRP.

Our solution for the SSP extends the recent approach of Arvelo et
al.~\cite{arvelo:aaai06} for QRPs that is based on the efficient 
enumeration of models for a propositional logic theory.
In our case, an input instance of the SSP is converted into an instance
of a QRP with preferences and constraints that is then translated
into a logical theory such that its best models are in correspondence
with the solutions of the SSP. These translations, from SSP to QRP to
logic, are performed efficiently, in (low) polynomial time, while the
best models are found using off-the-shelf SAT tools.
Thus, we are able to exploit the befits of modern SAT techniques such
as conflict-directed backtracking via clause learning and caching and
decomposition of common subproblems~\cite{beame:understanding} to
perform the necessary search in the combinatorial space of solutions.

@@

The semantic description of services is expressed following the
LAV paradigm in terms of mappings. These mappings describe the
services in terms of concepts in the domain ontology \cite{Ullman00}:
for each service $S_i$, there is a mapping that describes $S_i$
as a \emph{conjunctive query} on the concepts in the ontology that
also distinguish input and output attributes of the service.
For example, a service $S(x,y)$ that returns information about
flights originating at a given US city can be described with the view:
\[ S(\$x,y)\ \qrule\ \flight(x,y), \UScity(x)\,, \]
where $flight^2$ and $uscity^1$ are relational symbols in the
ontology. The symbol `\$' denotes that $x$ is an input attribute. 
The semantic interpretation of a mapping like this one enforce the following:
\begin{itemize}
\item the service represented by $S$ provides information in the
      form of tuples $(x,y)$,
\item the service is called with $x$ as input attribute and returns $(x,y)$,
\item each tuple $(x,y)$ returned by the service satisfies the rhs
      of the view; i.e., $\flight(x,y)$ and $\UScity(x)$, and
\item the views are not necessarily \emph{complete}; i.e., there may be
      other tuples $(x,y)$ that satisfy the rhs of the view but which
      are not available through $S$.
\end{itemize}

@@

\mcdsat\ does not provide support for constant symbols, yet incorporating
this functionality is straightforward.
Basically, one only has to track the unification of variables with
constants using new propositional symbols, and to propagate such unifications
transitively using implications in order to avoiding the unification of different
constant symbols.
This modification involve the addition of a small number of propositional
symbols and clauses to the CNF generated by \mcdsat.

@@

The subsumption relationships make the rewriting process more complex
as now one need to consider unification among predicate symbols of different
name and arity. Indeed, consider the following four concepts, where $a$,
$b$, $c$ and $d$ are constant symbols, and $x$, $y$ and $z$ are variables:
\begin{alignat*}{1}
P(b,y,z)\ &\orule\ R(a,y)\,, \\
R(a,y)\   &\orule\ T(c,y)\,, \\
P(d,x,z)\ &\orule\ M(a,x)\,, \\
M(a,x)\   &\orule\ N(d,x)\,,
\end{alignat*}
and the user request $Q(x,y)$ with services $S1$, $S2$ and $S3$:
\begin{alignat*}{1}
Q(x,y)\   &\qrule\ T(z,y),\,N(z,x)\,, \\
S_1(y)\   &\qrule\ R(a,y)\,, \\
S_2(x,z)\ &\qrule\ N(z,x)\,, \\
S_3(x,z)\ &\qrule\ P(d,x,z)\,.
\end{alignat*}
Then, the system must be able to infer that the query can be 
rewritten as $I(x,y) \qrule S_1(y),S_2(x,c)$ since $R(a,y)$
unifies with $T(z,y)$ producing the \emph{binding} $\{z=c\}$,
and $S_2(x,z)$ unifies with $N(z,x)$ and becomes $S_2(x,c)$
once the binding is propagated.
On the other hand, the system must also infer that $Q(x,y)$
cannot be rewritten as $I(x,y)\qrule S_1(y),S_3(x,z)$ because
$R(a,y)$ unifies with $T(z,y)$ with binding $\{z=c\}$,
$P(d,x,z)$ unifies with $N(z,x)$ with binding $\{z=d\}$,
and these two bindings are non-unifiable since constants
denote unique objects.

@@

The first benchmark consists of problems for air-travel queries.
Service views are of the form $V_i(x,y)\qrule \flight(x,y,\AL_i)$
where $\AL_i$ is a constant that denotes the name of an airline
and the view is assumed to return flights between two cities served
by the airline $\AL_i$. For the query, we consider a request to
find trips between Paris and New York with a number of stops.
The query returns the stops and has the form
\[ Q(x_1,\ldots,x_n)\ \qrule\ \flight(\PA,x_1,t),\,\flight(x_1,x_2,t),\,\ldots,\,\flight(x_n,\NY,t)\,. \]
Observe that the existentially quantified variable, $t$, is the same
for each flight meaning that it can only be unified with the same
constant; i.e., each leg of the flight is served by the same airline.
We solved several instances for this type of query with a number of stops
from 2 to 5 stops and a number of services from 10 to 100.
Fig.~\ref{fig:plot1}(a) shows the results of the compilations:
the vertical axis refers to the time in seconds in logarithmic
scale and the horizontal axis to the number of views in the benchmark.
The results show good performance since realistic instances of the
problem (sets of 100 airlines with 5-stop flights) can be compiled
in 328 seconds. The size in disk of the d-DNNF for 100 airlines and
5-stop flights is 3.4Mb from which the best model can be computed
in 0.29 seconds, and the enumeration of all models can be done
in 0.47 seconds.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth]{plots-v2/plot5}
\caption{Compilation times for experiments I and II for different
number of goals and different number of views.
Experiment II involves $n(n-1)/2$ preference of a problem with $n$ views.
The plots are in logarithmic scale, and the time is in seconds.}
\label{fig:plot1}
\end{figure}

@@

These existing solutions are scale to a number of abstract concepts.
In addition to scalability, our approach provides a more expressive
framework where services are semantically
described in terms of domain ontology concepts, user preferences
restrict the space of solutions, and ontology relationships augment
the space of possible solutions. Finally, our approach is sound and
complete.

>>>>

Existing Web infrastructures support the publication and access to a
tremendous amount of Web data sources, some of which can be labeled
and converted into Semantic Web Services by using existing annotation
tools like the one proposed by Ambite et al.\ \cite{AmbiteISWC09}.
Once a large dataset of Semantic Web Services become available, users
will require techniques to effectively select the services that meet
their requirements.
In order to achieve this goal, the services in the dataset must be
tagged with their functional and non-functional properties, and
the users' preferences and requirements must be formally described
as well.
In this paper, we extend an approach traditionally used in the area of
data integration to solve the problem of selecting the best services
that meet a user request, a problem that we call in this paper the
Service Selection Problem (SSP).

@@

As it is shown below, LAV descriptions are achieved with mappings
that define the services as conjunctive queries involving the generic
concepts in the ontology. Thus, every time that a service changes or
a new one becomes available, only a tiny fraction of the mappings
must be updated, usually just one mapping.
Likewise, user requests can be modeled as conjunctive queries over the
generic concepts in a way that the SSP can be cast as the problem
of rewriting a query in terms of a set of views, the so-called Query
Rewriting Problem (QRP) that is well-known in the area of data
integration~\cite{Chen05,JaudoinPRST05}, query optimization and data
maintenance~\cite{AfratiLU07,levy:bucket}, and for which several
scalable approaches had been
proposed~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
Furthermore, users' preferences and constraints on the set of the
selected services to serve a given request are supported by the
adoption of a simple yet expressive language for preferences. 
These preferences and constraints further refine and rank the
the set of valid rewritings of the posed query, in a way that
the best solution to the SSP corresponds to the best-ranked 
valid rewritings of the corresponding QRP.

Our solution for the SSP extends the recent approach of Arvelo et
al.~\cite{arvelo:aaai06} for QRPs that is based on the efficient 
enumeration of models for a propositional logic theory.
In our case, an input instance of the SSP is converted into an instance
of a QRP with preferences and constraints that is then translated
into a logical theory such that its best models are in correspondence
with the solutions of the SSP. These translations, from SSP to QRP to
logic, are performed efficiently, in (low) polynomial time, while the
best models are found using off-the-shelf SAT tools.
Thus, we are able to exploit the benefits of modern SAT techniques such
as conflict-directed backtracking via clause learning and caching and
decomposition of common subproblems~\cite{beame:understanding} to
perform the necessary search in the combinatorial space of solutions.

@@

The semantic description of services is expressed following the
LAV paradigm in terms of mappings. These mappings describe the
services in terms of concepts in the domain ontology \cite{Ullman00}:
for each service $S_i$, there is a mapping that describes $S_i$
as a \emph{conjunctive query} on the concepts in the ontology that
also distinguish input and output attributes of the service.
For example, a service $S(x,y)$ that returns information about
flights originating at a given US city can be described with the view:
\[ S(\$x,y)\ \qrule\ \flight(x,y), \UScity(x)\,, \]
where $flight^2$ and $uscity^1$ are relational symbols in the
ontology. The symbol `\$' denotes that $x$ is an input attribute. 
The semantic interpretation of a mapping like this one enforces the following:
\begin{itemize}
\item the service represented by $S$ provides information in the
      form of tuples $(x,y)$,
\item the service is called with $x$ as input attribute and returns $(x,y)$,
\item each tuple $(x,y)$ returned by the service satisfies the rhs
      of the view; i.e., $\flight(x,y)$ and $\UScity(x)$, and
\item the views are not necessarily \emph{complete}; i.e., there may be
      other tuples $(x,y)$ that satisfy the rhs of the view but which
      are not available through $S$.
\end{itemize}

@@

\mcdsat\ does not provide support for constant symbols, yet incorporating
this functionality is straightforward.
Basically, one only has to track the unification of variables with
constants using new propositional symbols, and to propagate such unifications
transitively using implications in order to avoiding the unification of different
constant symbols.
This modification involves the addition of a small number of propositional
symbols and clauses to the CNF generated by \mcdsat.

@@

The subsumption relationships make the rewriting process more complex
as now one needs to consider unification among predicate symbols of different
name and arity. Indeed, consider the following four concepts, where $a$,
$b$, $c$ and $d$ are constant symbols, and $x$, $y$ and $z$ are variables:
\begin{alignat*}{1}
P(b,y,z)\ &\orule\ R(a,y)\,, \\
R(a,y)\   &\orule\ T(c,y)\,, \\
P(d,x,z)\ &\orule\ M(a,x)\,, \\
M(a,x)\   &\orule\ N(d,x)\,,
\end{alignat*}
and the user request $Q(x,y)$ with services $S1$, $S2$ and $S3$:
\begin{alignat*}{1}
Q(x,y)\   &\qrule\ T(z,y),\,N(z,x)\,, \\
S_1(y)\   &\qrule\ R(a,y)\,, \\
S_2(x,z)\ &\qrule\ N(z,x)\,, \\
S_3(x,z)\ &\qrule\ P(d,x,z)\,.
\end{alignat*}
Then, the system must be able to infer that the query can be 
rewritten as $I(x,y) \qrule S_1(y),S_2(x,c)$ since $R(a,y)$
unifies with $T(z,y)$ producing the \emph{binding} $\{z=c\}$,
and $S_2(x,z)$ unifies with $N(z,x)$ and becomes $S_2(x,c)$
once the binding is propagated.
On the other hand, the system must also infer that $Q(x,y)$
cannot be rewritten as $I(x,y)\qrule S_1(y),S_3(x,z)$ because
$R(a,y)$ unifies with $T(z,y)$ with binding $\{z=c\}$,
$P(d,x,z)$ unifies with $N(z,x)$ with binding $\{z=d\}$,
and these two bindings are non-unifiable since constants
denote unique objects.

@@

The first benchmark consists of problems for air-travel queries.
Service views are of the form $V_i(x,y)\qrule \flight(x,y,\AL_i)$
where $\AL_i$ is a constant that denotes the name of an airline
and the view is assumed to return flights between two cities served
by the airline $\AL_i$. For the query, we consider a request to
find trips between Paris and New York with a number of stops.
The query returns the stops and has the form
\[ Q(x_1,\ldots,x_n)\ \qrule\ \flight(\PA,x_1,t),\,\flight(x_1,x_2,t),\,\ldots,\,\flight(x_n,\NY,t)\,. \]
Observe that the existentially quantified variable, $t$, is the same
for each flight meaning that it can only be unified with the same
constant; i.e., each leg of the flight is served by the same airline.
We solved several instances for this type of query with a number of stops
from 2 to 5 and a number of services from 10 to 100.
Fig.~\ref{fig:plot1}(a) shows the results of the compilations:
the vertical axis refers to the time in seconds in logarithmic
scale and the horizontal axis to the number of views in the benchmark.
The results show good performance since realistic instances of the
problem (sets of 100 airlines with 5-stop flights) can be compiled
in 328 seconds. The size in disk of the d-DNNF for 100 airlines and
5-stop flights is 3.4Mb from which the best model can be computed
in 0.29 seconds, and the enumeration of all models can be done
in 0.47 seconds.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth]{plots-v2/plot5}
\caption{Compilation times for experiments I and II for different
number of goals and different number of views.
Experiment II involves $n(n-1)/2$ preferences of a problem with $n$ views.
The plots are in logarithmic scale, and the time is in seconds.}
\label{fig:plot1}
\end{figure}

@@

These existing solutions scale to a number of abstract concepts.
In addition to scalability, our approach provides a more expressive
framework where services are semantically
described in terms of domain ontology concepts, user preferences
restrict the space of solutions, and ontology relationships augment
the space of possible solutions. Finally, our approach is sound and
complete.
